---
title: "Exercise 7"
editor_options: 
  chunk_output_type: console
format: docx
---

# Enrichment analysis

## Load Libraries
```{r}
#| warning: false
#| message: false
library(rGREAT)
library(TxDb.Mmusculus.UCSC.mm10.knownGene)
library(org.Mm.eg.db)
library(simplifyEnrichment)
library(flexclust)
```
```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE)
```

## OverlapMatrix
```{r}
overlap <- readRDS("output/overlap_anno2.rds")
```

## Running enrichment analysis
```{r}
gr <- overlap[overlap$ATAC_RNA == "Active Promoters"]
res <- great(gr, "BP", "mm10")

plotVolcano(res)
plotRegionGeneAssociations(res)
getRegionGeneAssociations(res)
```

```{r, eval=FALSE}
shinyReport(res)
```

```{r}
tb = getEnrichmentTable(res)
sig_go_ids = tb$id[tb$p_adjust < 0.05]
cl = simplifyGO(mat = sig_go_ids)
```

### Output
:::{.callout-important}
- The heatmap shows clear clustering of GO terms into two main clusters, indicating distinct functional groups.

- The GO terms on the right highlight biological processes relevant to neuronal and ion transport functions, such as transport, inorganic cation ion, exocytosis, synaptic, neurotransmitter, and behavior learning.

- The cluster with terms like cytosolic concentration, calcium ion suggests a calcium-related signaling or ion homeostasis process at play.

- The color gradient and similarity scores (up to ~0.6) indicate moderate similarity within clusters, which is reasonable for GO term groupings.
:::

### Interpretation
:::{.callout-important}
- E11.5 to E15.5 is a key window in neural development involving neuronal differentiation, synapse formation, and active signaling.

- Active promoters linked to ion transport, synaptic transmission, calcium signaling are expected to be regulated during this developmental stage.

- The clustering separating signaling and transport-related terms from behavior and learning terms also aligns with early neural circuit formation and functional maturation.
:::


## Question

**Perform enrichment analysis of any other category.**